---
title: "Clustering Homework"
output: html_notebook
---


```{r}
library(tidyverse)
library(janitor)
```

# 1

You have been provided some customer data from a shopping centre. Do some exploratory analysis on the data. Comment on findings.

```{r}
mall_data <- read_csv("data/mall_customers.csv") %>%
  clean_names()

glimpse(mall_data)
any(is.na(mall_data))
```

```{r}
mall_data %>%
  group_by(gender) %>%
  summarise(count = n(),
            avg_spend = mean(spending_score_1_100),
            avg_income = mean(annual_income_k))
```


```{r}
library(GGally)

ggpairs(mall_data)
```

```{r}
mall_data %>%
  ggplot() +
  aes(x = spending_score_1_100, y = annual_income_k, col = gender) +
  geom_point() +
  facet_wrap(~gender)
```

Patterns between spending score and annual income look consistent across both genders, so likely not a factor here. It seems like women have a higher spending score, but average income is less - however there are more female customers in the data than male.

The plot of the data shows potential for 5 distinct groups, one at each corner and one in the middle, but we should investigate further.

# 2 

We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers.

```{r}
# Remove non numeric columns, remove age as not needed, move id column to row labels, then scale.

mall_data_scale <- mall_data %>%
  column_to_rownames("customer_id") %>%
  select(-gender, -age) %>%
  mutate_if(is.numeric, scale)
  
```

## Perform k-means clustering and chose a value of k.

```{r}
# Start looking for best k-value

library(broom)

# Set min & max number of clusters want to look at 
max_k <- 20 

k_clusters <- tibble(k = 1:max_k) %>%
  mutate(
    kclust = map(k, ~ kmeans(mall_data_scale, .x, nstart = 25)), 
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, mall_data_scale)
  )

# Elbow plot
k_clusters %>%
  unnest(glanced) %>%
  ggplot() + 
  aes(x = k, y = tot.withinss) +
  geom_point() +
  geom_line() +
  scale_x_continuous(n.breaks = 20)
```

No clear elbow here apart from maybe 3 or 5, best to look at other plots.

```{r}
# Silhouette coefficient plot

library(factoextra)

fviz_nbclust(mall_data_scale,
             kmeans,
             method = "silhouette",
             nstart = 25)
```

Looks like 5 is the best in this one, one more plot.

```{r}
# Gap statistic plot

fviz_nbclust(mall_data_scale,
             kmeans,
             method = "gap_stat",
             nstart = 25)
```

For some reason dotted line by 1 here, but 5 looks better and also looks good on the other 2 plots.

# Visualise the clustering for your chosen value of k.

```{r}
# Generate K-means with 5 centers
mall_cluster <- kmeans(mall_data_scale, 
                       centers = 5, 
                       nstart = 25)

# Visualise clusters
augment(mall_cluster, mall_data) %>%
  ggplot() +
  aes(x = spending_score_1_100, y = annual_income_k, colour = .cluster) +
  geom_point(aes(colour = .cluster))
```

# Do you think the clustering seems a good fit for this data?

Yes! All of the clusters look like they fit well with the data.

# Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the ‘Segmentation & clustering intro’ lesson).

The "Window Shopper" - Low income & low spending score - perhaps younger customers hanging around at the mall, or people who are less well off scoping out their dream purchases.

The "Should Probably Try Amazon Prime" - Low income & high spending score - why are you spending so much at the mall? Probably suckeed in quite easily by a good deal or black friday sales.
