---
title: "Day 2 - rtweet"
output: html_notebook
---

```{r}
library(tidyverse)
```


# MVP

Question 1.
Load the code_clan_tweets.csv data. Find the number of rows, columns, and list all the variable names.

```{r}
tweet_data <- read_csv("data/code_clan_tweets.csv")

# view table
tweet_data

# view dimensions
dim(tweet_data)

# view variable names
names(tweet_data)
summary(tweet_data)
```


Question 2.
Find the total number of favourites (stored in favorite_count) that CodeClan tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet was quoted (see the is_quote variable).

```{r}
# filter tweet_data by is_quote being false. Then summarise the sum of favorite_count in total_favourites 
tweet_data %>%
  filter(is_quote == FALSE) %>%
  summarise(total_favourites = sum(favorite_count))
```


Question 3.
Summarise the mean number of retweets (retweet_count), split by the type of platform that was used to tweet (source). Again, omit quoted tweets from the mean.

```{r}
# Filter out quoted tweets, group by source then summarise average number of retweets per source
tweet_data %>%
  filter(is_quote == FALSE) %>%
  group_by(source) %>%
  summarise(mean_num_retweets = mean(retweet_count))
```


Question 4.
Count the total number of likes (i.e. total of favorite_count), split by media type, and arrange them from most likes to least. Assume that any tweets without a listed media type are of type “text”.

```{r}
# check media_types for distinct values
tweet_data %>%
  distinct(media_type)

# use if_else to mutate all NA values in media_type to "text"
# then group by media_type, summarise total favourites and arrange by descending order
tweet_data %>%
  mutate(media_type = if_else(is.na(media_type), "text", media_type)) %>%
  group_by(media_type) %>%
  summarise(total_likes = sum(favorite_count)) %>%
  arrange(desc(total_likes))

```


Question 5.
Find the mean number of characters that a CodeClan tweet contains. You can either find a way to count the text in the text variable, or use the variable display_text_width which contains a count. Checking both would be a good idea if possible to see if they match (but it’s not essential).

```{r}
# use str_count to compare length of tweet to the value in display_text_width
tweet_data %>%
  mutate(tweet_length = str_count(text)) %>%
  select(tweet_id, display_text_width, tweet_length)

# the counts definitely vary, having checked a random few manually I am satisfied that the count in the original dataset is not accurate, so I will use ours.
tweet_data %>%
  mutate(tweet_length = str_count(text)) %>%
  summarise(average_tweet_length = mean(tweet_length))
```



Question 6.
The code_clan_info.csv data file contains status_url along with other info. Load this in, and join it to the code_clan_tweets tibble, so that you have a status_url for each tweet. Decide which variable to join the tibbles on.

```{r}
# load in code_clan_info
code_clan_info <- read_csv("data/code_clan_info.csv")

# use inner_join so we end up with only tweets matched with the other table.
# join by variable "tweet_id"
tweet_data_joined <- tweet_data %>%
  inner_join(code_clan_info, by = "tweet_id")
```


Question 7.
From your new joined data, create a new tibble codeclan_hashtags containing only the tweet_id and convert the hashtags into lowercase for analysis. Keep only those tweets with hashtags.

```{r}
# select only tweet_id and hashtags columns, drop rows with no hashtags and convert to lower case.
codeclan_hashtags <- tweet_data_joined %>%
  select(tweet_id, hashtags) %>%
  drop_na() %>%
  mutate(hashtags = str_to_lower(hashtags))

codeclan_hashtags
```

#4 Extensions


Question 8.
Some tweets have more than one hashtag, denoted by the c( symbols at the start of the string. Use the str_sub() and str_detect() functions and your codeclan_hashtags tibble from above to find all the cases in which the hashtag string begins with characters c(.
 Hint Try a regex pattern "c\\(". The double backslash is necessary in R to mark a backslash as itself being escaped.
 
```{r}
# filter all rows in hashtags beginning with "c("
# str_sub to remove c( and ) from start and end
codeclan_hashtags %>%
  filter(str_detect(hashtags, "c\\(")) %>%
  mutate(hashtags = str_sub(hashtags, start = 3, end = -2))
```
 


Question 9.
Use the str_detect() function to find all tweets with text that mentions “Edinburgh”, and count how many such tweets there are.
 Hint You may need to make sure your tweets are all the same case before searching.

```{r}
# filter with str_detect() and regex(ignore_case)
tweet_data %>%
  select(text) %>%
  filter(str_detect(text, regex("Edinburgh", ignore_case = TRUE))) %>%
  count()
  
# or mutate to lowercase using str_to_lower() and filter with lowercase "edinburgh"
tweet_data %>%
  select(text) %>%
  mutate(text = str_to_lower(text)) %>%
  filter(str_detect(text, "edinburgh")) %>%
  count()

# both return 33
```


Question 10.
Use str_extract_all() with a regular expression to find out which Twitter users CodeClan have been tweeting.
 Hint Think about all the different characters which could be in a Twitter username, do a bit of online research

```{r}
# regular expression is: and @, followed by all alphanumeric including underscores (this is all twitter permits) then "+" for one or more character.
# must unnest the usernames as str_extract_all will output a list per row, then show all distinct usernames.
tweet_data %>%
  mutate(usernames = str_extract_all(text ,"@[\\w]+")) %>%
  unnest(usernames) %>%
  distinct(usernames)
```
