---
title: "Multiple Regression Homework"
output: html_notebook
---

```{r}
library(tidyverse)
```

# MVP

## 1. Load the diamonds.csv data set and undertake an initial exploration of the data. You will find a description of the meanings of the variables on the relevant Kaggle page

```{r}
diamonds <- read_csv("data/diamonds.csv")

glimpse(diamonds)
diamonds
```

## 2. We expect the carat of the diamonds to be strong correlated with the physical dimensions x, y and z. Use ggpairs() to investigate correlations between these four variables.
```{r}
library(GGally)

diamonds %>%
  select(carat, c(x:z)) %>%
  ggpairs()
```

## 3. So, we do find significant correlations. Let’s drop columns x, y and z from the dataset, in preparation to use only carat going forward.

```{r}
diamonds_trimmed <- diamonds %>%
  select(-c(1,x:z))
```

## 4. We are interested in developing a regression model for the price of a diamond in terms of the possible predictor variables in the dataset.

### i. Use ggpairs() to investigate correlations between price and the predictors (this may take a while to run, don’t worry, make coffee or something).

```{r}
ggpairs(diamonds_trimmed)
```


### ii. Perform further ggplot visualisations of any significant correlations you find.

```{r}
diamonds_trimmed %>%
  ggplot() +
  aes(x = price, y = cut) +
  geom_boxplot()
```

```{r}
diamonds_trimmed %>%
  ggplot() +
  aes(x = price, y = color) +
  geom_boxplot()
```

```{r}
diamonds_trimmed %>%
  ggplot() +
  aes(x = price, y = clarity) +
  geom_boxplot()
```

## 5. Shortly we may try a regression fit using one or more of the categorical predictors cut, clarity and color, so let’s investigate these predictors:

### i. Investigate the factor levels of these predictors. How many dummy variables do you expect for each of them?

```{r}
diamonds_trimmed %>%
  distinct(cut)

# 5 distinct values, so 4 dummy variables
```
```{r}
diamonds_trimmed %>%
  distinct(color)

# 7 distinct values, so 6 dummy variables
```

```{r}
diamonds_trimmed %>%
  distinct(clarity)

# 8 distinct values, so 7 dummy variables
```

### ii. Use the dummy_cols() function in the fastDummies package to generate dummies for these predictors and check the number of dummies in each case.

```{r}
library(fastDummies)
diamonds_dummy <- diamonds %>%
  dummy_cols(select_columns = c("cut", "color", "clarity"),
             remove_selected_columns = TRUE)

diamonds_dummy
```

# 6. Start with simple linear regression. Regress price on carat and check the regression diagnostics.

```{r}
model_1 <- lm(price ~ carat, data = diamonds_trimmed)

autoplot(model_1)
summary(model_1)
```

# 7. Add another predictor of your choice. Check your assumptions, diagnostics, and interpret the model.

```{r}
model_2 <- lm(price ~ carat + cut, data = diamonds_trimmed)

autoplot(model_2)
summary(model_2)
```

