---
title: "Extra Practice Homework"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
library(janitor)
```

```{r}
disaster_data <- read_csv(here("raw_data/disasters_with_errors.csv"))
```

```{r}
summary(disaster_data)

disaster_data %>%
  summarise(across(everything(), ~sum(is.na(.x))))

# remove duplicate rows
disaster_distinct <- disaster_data %>%
  select(-id) %>%
  distinct()
```



```{r}
# Check iso column for misspelling/non 3 letter codes
disaster_distinct %>%
  select(iso) %>%
  distinct(iso) %>%
  filter(str_detect(iso, "[\\w]{4,}"))

# Only china needs fixed
disaster_iso_fixed <- disaster_distinct %>%
  mutate(iso = if_else(iso == "CHINA", "CHN", iso))
```

```{r}
# Check disaster_type for misspellings
disaster_iso_fixed %>%
  select(disaster_type) %>%
  distinct(disaster_type) %>%
  arrange(disaster_type)

disaster_types_fixed <- disaster_iso_fixed %>%
  mutate(disaster_type = case_when(
    disaster_type == "Droght" ~ "Drought",
    disaster_type == "Miscelaneous accident" ~ "Miscellaneous accident",
    TRUE ~ disaster_type)
  )
```

```{r}
# drop all rows where there is no data from total_deaths to total_damage
disaster_na_removed <- disaster_types_fixed %>%
  # drop total_damage as we don't need it
  select(-total_damage) %>%
  filter(!if_all(total_deaths:total_affected, ~is.na(.x)))
```


```{r}
# impute total_affected column with the sum of all other columns if it is NA.
disaster_data_clean <- disaster_na_removed %>%
  rowwise() %>%
  mutate(total_affected = case_when(
      is.na(total_affected) ~ sum(total_deaths, affected, injured, homeless, na.rm = TRUE),
      TRUE ~ total_affected
    ),
    # replace NA in total_deaths and affected with 0
    total_deaths = case_when(
      is.na(total_deaths) ~ 0,
      TRUE ~ total_deaths
    ), 
    affected = case_when(
      is.na(affected) ~ 0,
      TRUE ~ affected
    )
  ) %>%
  ungroup() %>%
  select(-injured, -homeless) %>%
  # remove outliers
  filter(total_deaths >= 10 | total_affected >= 100)

```

### For a disaster to be entered into the database at least one of the following criteria must be fulfilled:

#### Ten (10) or more people reported killed
#### Hundred (100) or more people reported affected
#### https://www.emdat.be/explanatory-notes

# Questions

## 1. What was the greatest disaster of all time by total deaths?
```{r}
disaster_data_clean %>%
  select(year, disaster_type, country_name, total_deaths) %>%
  arrange(desc(total_deaths)) %>%
  slice_max(total_deaths, n = 1, with_ties = TRUE)
```

## 2. Which country has had the highest total people affected by disasters?
```{r}
disaster_data_clean %>%
  group_by(country_name) %>%
  summarise(all_affected = sum(total_affected)) %>%
  slice_max(all_affected, n = 1, with_ties = TRUE)

```

## 3. What are the top 5 countries for total number of disasters?
```{r}
disaster_data_clean %>%
  group_by(country_name) %>%
  summarise(total_number_disasters = sum(occurrence)) %>%
  slice_max(total_number_disasters, n = 1, with_ties = TRUE)
```

## 4. What is the greatest disaster in the last 5 years of the data, by total deaths?
```{r}
disaster_data_clean %>%
  filter(year > max(year) - 5) %>%
  slice_max(total_deaths, n = 1, with_ties = TRUE)
```

## 5. What were the top 10 years that had the most people dying from earthquakes?
```{r}
disaster_data_clean %>%
  group_by(year) %>%
  filter(disaster_type == "Earthquake") %>%
  summarise(total_earthquake_deaths = sum(total_deaths)) %>%
  arrange(desc(total_earthquake_deaths)) %>%
  head(10)
```

