---
title: "Manual model development"
output: html_notebook
---

# MVP

## 1.

Tidy up the data ready for regression:

* You might like to think about removing some or all of `date`, `id`, `sqft_living15`, `sqft_lot15` and `zipcode` (`lat` and `long` provide a better measure of location in any event).
* Have a think about how to treat `waterfront`. Should we convert its type?
* We converted `yr_renovated` into a `renovated` logical variable, indicating whether the property had ever been renovated. You may wish to do the same.
* Have a think about how to treat `condition` and `grade`? Are they interval or categorical ordinal data types?

```{r}
library(tidyverse)
library(modelr)
library(GGally)
library(ggfortify)
```

```{r}
house_data <- read_csv("data/kc_house_data.csv")

house_data_trim <- house_data %>%
  select(-c("date", "id", "sqft_living15", "sqft_lot15", "zipcode")) %>%
  mutate(waterfront = if_else(waterfront == 1, TRUE, FALSE),
         yr_renovated = if_else(yr_renovated != 0, TRUE, FALSE),
         condition = case_when(
           condition == 1 ~ "Poor",
           condition == 2 ~ "Fair",
           condition == 3 ~ "Good",
           condition == 4 ~ "Very Good",
           condition == 5 ~ "Excellent"
         ),
         grade = case_when(
           grade <= 3 ~ "Poor",
           grade < 11 ~ "Average",
           TRUE ~ "Excellent"
         )) %>%
  rename(renovated = yr_renovated)
```

## 2.
Check for aliased variables using the alias() function (this takes in a formula object and a data set). [Hint - formula price ~ . says ‘price varying with all predictors’, this is a suitable input to alias()]. Remove variables that lead to an alias. Check the ‘Elements of multiple regression’ lesson for a dropdown containing further information on finding aliased variables in a dataset.

```{r}
alias(price ~ ., data = house_data_trim)
```

```{r}
# Remove sqft_basement 
house_data_trim <- house_data_trim %>%
  select(-sqft_basement)
```

## 3.
Systematically build a regression model containing up to four main effects (remember, a main effect is just a single predictor with coefficient), testing the regression diagnostics as you go * splitting datasets into numeric and non-numeric columns might help ggpairs() run in manageable time, although you will need to add either a price or resid column to the non-numeric dataframe in order to see its correlations with the non-numeric predictors.

```{r}
ggpairs(house_data_trim)
```

Strongest correlation between price and sqft_living (0.702)

```{r}
model_1 <- lm(formula = price ~ sqft_living, data = house_data_trim)

summary(model_1)
```

```{r}
autoplot(model_1)
```

The r squared value for the model including sqft_living indicates 49% of the variation of price can be explained. The diagnostic criteria have been met. (maybe not Scale-Location?)

## Second predictor

```{r}
house_remaining_resid <- house_data_trim %>%
  add_residuals(model_1) %>%
  select(-c("price", "sqft_living"))

house_remaining_resid %>%
  ggpairs()
```

Looks to be variation in both condition and grade, will test these two for the model

```{r}
model_2a <- lm(formula = price ~ sqft_living + condition, data = house_data_trim)

summary(model_2a)
```


```{r}
model_2b <- lm(formula = price ~ sqft_living + grade, data = house_data_trim)

summary(model_2b)
```

Model 2b, sqft_living + grade has the higher r2 value.

```{r}
autoplot(model_2b)
```

# Third predictor

```{r}
house_remaining_resid <- house_data_trim %>%
  add_residuals(model_2b) %>%
  select(-c("price", "sqft_living", "grade"))

house_remaining_resid %>%
  ggpairs()
```

```{r}
model_3a <- lm(formula = price ~ sqft_living + grade + lat, data = house_data_trim)

summary(model_3a)
```

```{r}
autoplot(model_3a)
```

```{r}
anova(model_2b, model_3a)
```

# Fourth predictor
```{r}
house_remaining_resid <- house_data_trim %>%
  add_residuals(model_3a) %>%
  select(-c("price", "sqft_living", "grade", "lat"))

house_remaining_resid %>%
  ggpairs()
```

```{r}
model_4a <- lm(formula = price ~ sqft_living + grade + lat + view, data = house_data_trim)

summary(model_4a)
```

```{r}
autoplot(model_4a)
```

