---
title: "Advanced ggplot homework"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
library(CodeClanData)

revenue_breakdown <- qb_revenue_breakdown
monthly_sales <- qb_monthly_sales
competitors <- qb_competitors
device_data <- qb_device_data
```

# MVP
## Part 1

### 1. Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

```{r}
ggplot(revenue_breakdown) +
  aes(x = Year, y = Revenue, fill = Product) +
  geom_col() 
```

### 2. Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).

```{r}
monthly_sales %>%
  filter(Cashflow == "Kwikbit Unit 1 Sales") %>%
  ggplot() +
  aes(x = Date, y = Amount, colour = Cashflow) +
  geom_line() +
  scale_x_date(breaks = "1 month") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

### 3. Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.

```{r}
monthly_sales %>%
  filter(Cashflow != "Kwikbit Unit 1 Sales") %>%
  ggplot() +
  aes(x = Date, y = Amount, colour = Cashflow) +
  geom_line() +
  scale_x_date(breaks = "1 month") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

### 4. Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.

```{r}
competitors %>%
  ggplot() +
  aes(x = Year, y = Revenue, fill = Company) +
  geom_area() 
```

### 5. Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
competitors %>%
  ggplot() +
  aes(x = Year, y = Revenue, colour = Company) +
  geom_line() +
  geom_point()
```

### 6. Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in qb_device_data.

```{r}
device_data %>%
  # First total up number of steps per hour for each device/user combination
  group_by(device, id, hours) %>%
  summarise(counts = sum(counts)) %>%
  ggplot() +
    aes(x = hours, y = counts, colour = device) +
    geom_line() +
    facet_grid(id ~ device) +
    theme_minimal() +
    theme(legend.position = "none")
```


## Part 2
Take the plots that you produced in part one and now polish them by:

Adding appropriate labels

Changing the scales and coordinates when appropriate.

Applying a unified theme, which is described below:
Graphs should have white backgrounds, and use colour sparingly.
There should be faint grid lines.
Font sizes should be ~12pt although titles should be slightly larger and axis labels can be slightly smaller.
All plots should use colours from the following company colour scheme.
```{r}
# Set company colour scheme
col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")
```

### 1,
```{r}
ggplot(revenue_breakdown) +
  aes(x = Year, y = Revenue, fill = Product) +
  geom_col() +
  # add labels
  labs(
    x = "\nYear",
    y = "Revenue\n",
    title = "\nTotal Yearly Revenue",
    subtitle = "Split by product type\n",
    fill = "Product Type"
  ) +
  # add theme elements
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 10, face = "bold"),
    axis.text = element_text(size = 8),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "light grey", size = 0.1)
  ) +
  scale_fill_manual(values = col_scheme)
```

### 2.
```{r}
monthly_sales %>%
  filter(Cashflow == "Kwikbit Unit 1 Sales") %>%
  ggplot() +
    aes(x = Date, y = Amount, colour = Cashflow) +
    geom_line() +
    # change x scale to monthly, in nice date format
    scale_x_date(breaks = "1 month", labels = scales::date_format("%b-%Y")) +
    # add labels
    labs(
      x = "\nDate",
      y = "Sales\n",
      title = "\nKwikbit Unit 1 Sales",
      subtitle = "Over last 12 months\n"
    ) +
    # add theme elements
    theme(
      # remove legend
      legend.position = "none",
      # rotate x axis 45 degrees and nudge down 
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(size = 16, face = "bold"),
      plot.subtitle = element_text(size = 12),
      axis.title = element_text(size = 10, face = "bold"),
      axis.text = element_text(size = 8),
      panel.background = element_rect(fill = "white"),
      panel.grid = element_line(colour = "light grey", size = 0.1)
    ) + 
    scale_colour_manual(values = col_scheme)

```

### 3.
```{r}
monthly_sales %>%
  filter(Cashflow != "Kwikbit Unit 1 Sales") %>%
  ggplot() +
    aes(x = Date, y = Amount, colour = Cashflow) +
    geom_line() +
    # add labels
    labs(
      x = "\nDate",
      y = "Amount\n",
      colour = "Cash flow",
      title = "\nTotal Revenue and Costs",
      subtitle = "Over last 12 months\n"
    ) +
    # change y axis breaks
    scale_y_continuous(breaks = c(1500, 1750, 2000, 2250, 2500, 2750)) +
    # change x scale to monthly, in nice date format
    scale_x_date(breaks = "1 month", labels = scales::date_format("%b-%Y")) +
    # add theme elements
    theme(
      # rotate x axis 45 degrees and nudge down 
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(size = 16, face = "bold"),
      plot.subtitle = element_text(size = 12),
      legend.title = element_text(size = 14, face = "bold"),
      legend.text = element_text(size = 12),
      axis.title = element_text(size = 10, face = "bold"),
      axis.text = element_text(size = 8),
      panel.background = element_rect(fill = "white"),
      panel.grid = element_line(colour = "light grey", size = 0.1)
    ) +
    scale_colour_manual(values = col_scheme)
```

### 4. 

```{r}
competitors %>%
  ggplot() +
    aes(x = Year, y = Revenue, fill = Company) +
    geom_area() +
    # add labels
    labs(
      x = "\nYear",
      y = "Revenue\n",
      fill = "Business",
      title = "\nSales of Personal Fitness Trackers",
      subtitle = "Over last 5 years\n"
    ) +
    # Change y axis from scientific notation
    scale_y_continuous(labels = scales::comma) +
    # add theme elements
    theme(
      plot.title = element_text(size = 16, face = "bold"),
      plot.subtitle = element_text(size = 12),
      legend.title = element_text(size = 14, face = "bold"),
      legend.text = element_text(size = 12),
      axis.title = element_text(size = 10, face = "bold"),
      axis.text = element_text(size = 8),
      panel.background = element_rect(fill = "white"),
      panel.grid = element_line(colour = "light grey", size = 0.1)
    ) +
    scale_fill_manual(values = col_scheme)
```

### 5. 
```{r}
competitors %>%
  ggplot() +
    aes(x = Year, y = Revenue, colour = Company) +
    geom_line() +
    geom_point() + 
    # add labels
    labs(
      x = "\nYear",
      y = "Revenue\n",
      colour = "Business",
      title = "\nSales of Personal Fitness Trackers",
      subtitle = "Over last 5 years\n"
    ) +
    # Change y axis from scientific notation
    scale_y_continuous(labels = scales::comma) +
    # add theme elements
    theme(
      plot.title = element_text(size = 16, face = "bold"),
      plot.subtitle = element_text(size = 12),
      legend.title = element_text(size = 14, face = "bold"),
      legend.text = element_text(size = 12),
      axis.title = element_text(size = 10, face = "bold"),
      axis.text = element_text(size = 8),
      panel.background = element_rect(fill = "white"),
      panel.grid = element_line(colour = "light grey", size = 0.1)
    ) +
    scale_colour_manual(values = col_scheme)
```

### 6.
```{r}
device_data %>%
  # First total up number of steps per hour for each device/user combination
  group_by(device, id, hours) %>%
  summarise(counts = sum(counts)) %>%
  # Tidy up names
  mutate(
    device = case_when(
      device == "qikbit1" ~ "Qikbit 1",
      device == "competitor1" ~ "Competitor 1",
      device == "competitor2" ~ "Competitor 2",
      device == "competitor3" ~ "Competitor 3"
    ),
    id = case_when(
      id == "person1" ~ "User 1",
      id == "person2" ~ "User 2",
      id == "person3" ~ "User 3",
      id == "person4" ~ "User 4",
      id == "person5" ~ "User 5"
    ),
    # Order devices as factor for graph
    device = factor(device, levels = c("Qikbit 1", "Competitor 1", "Competitor 2", "Competitor 3"))
    ) %>%
  ggplot() +
    aes(x = hours, y = counts, colour = device) +
    geom_line() +
    # facet by user and device
    facet_grid(id ~ device) +
    # add labels
    labs(
      x = "\nHours",
      y = "Number of Steps\n",
      fill = "Business",
      title = "\nQikbit vs Competitors' Devices",
      subtitle = "Step count data from 5 users\n"
    ) +
    # change x axis breaks
    scale_x_continuous(breaks = c(5, 8, 11, 14, 17, 20, 23)) +
    # add theme elements
    theme(
      # remove legend
      legend.position = "none",
      plot.title = element_text(size = 16, face = "bold"),
      plot.subtitle = element_text(size = 12),
      axis.title = element_text(size = 10, face = "bold"),
      axis.text = element_text(size = 8),
      panel.background = element_rect(fill = "white"),
      panel.grid = element_line(colour = "light grey", size = 0.1),
      strip.text = element_text(size = 8, face = "bold")
    ) +
    scale_colour_manual(values = col_scheme)
```


# Extension
Try out different ways of plotting these three datasets. Which graphs do you think are particularly effective at showing QikBit’s success?

### Q1 but bar chart with dodge
```{r}
ggplot(revenue_breakdown) +
  aes(x = Year, y = Revenue, fill = Product) +
  geom_col(position = "dodge") +
  # add labels
  labs(
    x = "\nYear",
    y = "Revenue\n",
    title = "\nTotal Yearly Revenue",
    subtitle = "Split by product type\n",
    fill = "Product Type"
  ) +
  # add theme elements
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 10, face = "bold"),
    axis.text = element_text(size = 8),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "light grey", size = 0.1)
  ) +
  scale_fill_manual(values = col_scheme)

```

### Q2 but bar chart
```{r}
monthly_sales %>%
  filter(Cashflow == "Kwikbit Unit 1 Sales") %>%
  ggplot() +
    aes(x = Date, y = Amount, fill = Cashflow) +
    geom_col() +
    # change x scale to monthly, in nice date format
    scale_x_date(breaks = "1 month", labels = scales::date_format("%b-%Y")) +
    # add labels
    labs(
      x = "\nDate",
      y = "Sales\n",
      title = "\nKwikbit Unit 1 Sales",
      subtitle = "Over last 12 months\n"
    ) +
    # add theme elements
    theme(
      # remove legend
      legend.position = "none",
      # rotate x axis 45 degrees and nudge down 
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(size = 16, face = "bold"),
      plot.subtitle = element_text(size = 12),
      axis.title = element_text(size = 10, face = "bold"),
      axis.text = element_text(size = 8),
      panel.background = element_rect(fill = "white"),
      panel.grid = element_line(colour = "light grey", size = 0.1)
    ) + 
    scale_fill_manual(values = col_scheme)

```

### Q4/5 but bar chart with dodge
```{r}
competitors %>%
  ggplot() +
    aes(x = Year, y = Revenue, fill = Company) +
    geom_col(position = "dodge") +
    # add labels
    labs(
      x = "\nYear",
      y = "Revenue\n",
      colour = "Business",
      title = "\nSales of Personal Fitness Trackers",
      subtitle = "Over last 5 years\n"
    ) +
    # Change y axis from scientific notation
    scale_y_continuous(labels = scales::comma) +
    # add theme elements
    theme(
      plot.title = element_text(size = 16, face = "bold"),
      plot.subtitle = element_text(size = 12),
      legend.title = element_text(size = 14, face = "bold"),
      legend.text = element_text(size = 12),
      axis.title = element_text(size = 10, face = "bold"),
      axis.text = element_text(size = 8),
      panel.background = element_rect(fill = "white"),
      panel.grid = element_line(colour = "light grey", size = 0.1)
    ) +
    scale_fill_manual(values = col_scheme)
```

### Q6 but Histogram
```{r}
device_data %>%
  # First total up number of steps per hour for each device/user combination
  group_by(device, id, hours) %>%
  summarise(counts = sum(counts)) %>%
  # Tidy up names
  mutate(
    device = case_when(
      device == "qikbit1" ~ "Qikbit 1",
      device == "competitor1" ~ "Competitor 1",
      device == "competitor2" ~ "Competitor 2",
      device == "competitor3" ~ "Competitor 3"
    ),
    id = case_when(
      id == "person1" ~ "User 1",
      id == "person2" ~ "User 2",
      id == "person3" ~ "User 3",
      id == "person4" ~ "User 4",
      id == "person5" ~ "User 5"
    ),
    # Order devices as factor for graph
    device = factor(device, levels = c("Qikbit 1", "Competitor 1", "Competitor 2", "Competitor 3"))
    ) %>%
  ggplot() +
    aes(x = hours, y = counts, fill = device) +
    geom_histogram(stat = "identity") +
    # facet by user and device
    facet_grid(id ~ device) +
    # add labels
    labs(
      x = "\nHours",
      y = "Number of Steps\n",
      fill = "Business",
      title = "\nQikbit vs Competitors' Devices",
      subtitle = "Step count data from 5 users\n"
    ) +
    # change x axis breaks
    scale_x_continuous(breaks = c(5, 8, 11, 14, 17, 20, 23)) +
    # add theme elements
    theme(
      # remove legend
      legend.position = "none",
      plot.title = element_text(size = 16, face = "bold"),
      plot.subtitle = element_text(size = 12),
      axis.title = element_text(size = 10, face = "bold"),
      axis.text = element_text(size = 8),
      panel.background = element_rect(fill = "white"),
      panel.grid = element_line(colour = "light grey", size = 0.1),
      strip.text = element_text(size = 8, face = "bold")
    ) +
    scale_fill_manual(values = col_scheme)

```