---
title: "Day 2 Homework: Starbucks Calories"
output: html_notebook
---

# Question 1.

Load in the dataset and have a look at it. What are the dimensions, variable types, variable names, etc.?

```{r}
# Load tidyverse package
library(tidyverse)

# Load in dataset
all_calories <- read.csv("data/starbucks_drinkMenu_expanded.csv")
```

```{r}
# Check data for dimensions, variable types, names, etc.
head(all_calories)
dim(all_calories)
names(all_calories)
glimpse(all_calories)
```
We can see there are 242 observations of 18 variables.
The variables are relating to drinks available at Starbucks, such as Beverage, Beverage_prep, Calories, and various other nutritional information.
There is a mix of factors and numerical data (integers and doubles) in this dataset.



# Question 2.

Let’s first investigate the calories of different drinks. Select the variables Beverage_category, Beverage, Beverage prep and Calories from your data. Since we are interested in the calorie content, check if there are any NA values in the data, and drop them if there are.

```{r}
# Select variables
calorie_info <- select(all_calories, Beverage_category, Beverage, Beverage_prep, Calories)
calorie_info
```

```{r}
# Check calories for NA values
is.na(calorie_info$Calories)
```
There are no NA values in this data, but this is how you'd filter them out anyway.

```{r}
# Filter out NA values
calories_filtered <- filter(calorie_info, !is.na(Calories))
calories_filtered
```



# Question 3.

Create a new variable (column) called calorie_diff, which stores the difference between 135 calories (135 calories = 10 cubes of sugar!) and the calories in each drink. (hint: you’ll want to subtract 135 from the calories to see which drink have more or less than 10 cups of sugar).

```{r}
# Create calorie_diff variable
calories_difference <- mutate(calories_filtered, calorie_diff = Calories - 135)
calories_difference
```


# Question 4.

Summarise the mean number of calories in each beverage_category. Which 3 category of drinks have the most calories? Which 3 drinks have the least? Write a small summary of your findings.

```{r}
# Group by beverage_category
beverages_grouped <- group_by(calories_difference, Beverage_category)

# Summarise mean calories per beverage_category
calorie_stats <- summarise(beverages_grouped, mean_calories = mean(Calories))
calorie_stats

# Arrange by descending order
calorie_stats_arranged <- arrange(calorie_stats, desc(mean_calories))
calorie_stats_arranged

# Show Top 3
head(calorie_stats_arranged, 3)

# Show Bottom 3
tail(calorie_stats_arranged, 3)
```
```{r}
# Or use pipe
calorie_stats_arranged <- calories_difference %>%
  group_by(Beverage_category) %>%
  summarise(mean_calories = mean(Calories)) %>%
  arrange(desc(mean_calories))
  
calorie_stats_arranged

# Show Top 3
head(calorie_stats_arranged, 3)

# Show Bottom 3
tail(calorie_stats_arranged, 3)
```


The three beverage categories with the highest calories are Smoothies, Frappuccinos and Signature Espresso Drinks.The three lowest are Classic Espresso Drinks, Shaken Iced Beverages and Coffee.

Both the largest and smallest calorie counts, Smoothies and Coffee, have some of the smallest sample sizes of their categories in this dataset. This could show that using the beverage category may not be the best for comparison.

# Question 5.

Let’s look at this a different way. What is the average number of calories in each Beverage_prep type?

```{r}
# Pipe function to group beverage_prep, summarise average calories, and arrange descending by calories
average_prep_calories <- calories_filtered %>%
  group_by(Beverage_prep) %>%
  summarise(average_calories = mean(Calories)) %>%
  arrange(desc(average_calories))

average_prep_calories
```

