---
title: "Day 2 Homework: Starbucks Calories"
output: html_notebook
---

# Question 1.

Load in the dataset and have a look at it. What are the dimensions, variable types, variable names, etc.?

```{r}
# Load tidyverse package
library(tidyverse)

# Load in dataset
all_calories <- read.csv("data/starbucks_drinkMenu_expanded.csv")
```

```{r}
# Check data for dimensions, variable types, names, etc.
head(all_calories)
dim(all_calories)
names(all_calories)
glimpse(all_calories)
```
We can see there are 242 observations of 18 variables.
The variables are relating to drinks available at Starbucks, such as Beverage, Beverage_prep, Calories, and various other nutritional information.
There is a mix of factors and numerical data (integers and doubles) in this dataset.



# Question 2.

Let’s first investigate the calories of different drinks. Select the variables Beverage_category, Beverage, Beverage prep and Calories from your data. Since we are interested in the calorie content, check if there are any NA values in the data, and drop them if there are.

```{r}
# Select variables
calorie_info <- select(all_calories, Beverage_category, Beverage, Beverage_prep, Calories)
calorie_info
```

```{r}
# Check calories for NA values
is.na(calorie_info$Calories)
```
There are no NA values in this data, but this is how you'd filter them out anyway.

```{r}
# Filter out NA values
calories_filtered <- filter(calorie_info, !is.na(Calories))
calories_filtered
```

