---
title: "Probability homework"
output: html_notebook
---

```{r}
library(prob)
library(gtools)
library(tidyverse)
```

# MVP - Titanic
```{r}
titanic_data <- read_csv("data/titanic.csv")

# check distinct values in class column
titanic_data %>%
  distinct(class)

# use str_detect with regex pattern to create new column "crew" with true/false values
titanic_crew <- titanic_data %>%
  mutate(crew = str_detect(class, "crew|staff"))
```


```{r}
# probability of being crew and survived
p_crew_and_survived <- titanic_crew %>%
  filter(crew == TRUE & survived == "yes") %>%
  summarise(p = n() / nrow(titanic_crew))

p_crew_and_survived
```
P(survived | crew) = 0.096
```{r}
# probability of being passenger and survived
p_passenger_and_survived <- titanic_crew %>%
  filter(crew == FALSE & survived == "yes") %>%
  summarise(p = n() / nrow(titanic_crew))

p_passenger_and_survived
```
p(survived | passenger) = 0.226

The probability of surviving as crew is 0.096, and surviving as a passenger is 0.226, meaning passengers were more than twice as likely to survive.

# MVP / Extension - Probability problems

## 1. If I roll a six-sided dice once, what is the probability of getting a number less than 5?

```{r}
dice <- rolldie(1)

p_less_than_5 <- dice %>%
  filter(X1 < 5) %>%
  summarise(p = n() / nrow(dice))

p_less_than_5
```
P(roll less than 5) = 0.66 or 2/3

## 2. How many outcomes are there in the sample space of an experiment in which we toss three dice?

```{r}
three_die <- rolldie(3)

n_outcomes_three_die <- three_die %>%
  summarise(n())

n_outcomes_three_die
```
216 possible outcomes of rolling three die.

## 3. In an experiment where we roll four dice:

```{r}
four_die <- rolldie(4)
```

### how many outcomes are in the event where the total of the four dice equals 5, 6 or 7?
```{r}
four_die_totals <- four_die %>%
  mutate(total = rowSums(.))

four_die_totals %>%
  filter(total %in% c(5, 6, 7)) %>%
  summarise(n())
```
There are 34 outcomes where the total of four die is 5, 6 or 7

### what is the probability of this event?
```{r}
four_die_totals %>%
  filter(total %in% c(5, 6, 7)) %>%
  summarise(p = n() / nrow(four_die_totals))
```
P(total of four die is 5, 6 or 7) = 0.026 or around 1/38

### what assumptions do we apply in answering part ii?

Not sure how to answer this one! Maybe that the order of the dice don't matter?


## 4. Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

## Hint: This is an arrangement where order matters. Of all the possible arrangements, how many are correct?

```{r}
heights <- c(1.6, 1.7, 1.8, 1.9, 2.0)

heights_orders <- as_tibble(
  permutations(
    n = 5,
    r = 5,
    v = heights,
    repeats.allowed = FALSE
  )
)

nrow(heights_orders)
```

There are 120 possible combinations of the 5 people sitting in the 5 chairs. We know that they are all differing height so there is only one possible outcome where they are sitting in order of descending height, left to right. This means that:
P(desc height left to right) = 1/120

## 5. I roll six dice:
```{r}
six_die <- rolldie(6)
```

### what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?
```{r}
six_die %>%
  filter(isin(., c(6, 5, 4, 3, 2, 1), ordered = TRUE)) %>%
  summarise(p = n() / nrow(six_die))

```
P(6, 5, 4, 3, 2 and 1 in order) = 2.14e-05 or 1/46656

### what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?
### [Hint have a look at the isin() function in the prob package. For example, to filter all rows containing 6, 1, 2 in that order, you could use isin(., c(6, 1, 2), ordered = TRUE)]
```{r}
six_die %>%
  filter(isin(., c(6, 5, 4, 3, 2, 1), ordered = FALSE)) %>%
  summarise(p = n() / nrow(six_die))
```
P(one each of 6, 5, 4, 3, 2 and 1) = 0.015 or 720/46656 or 5/324

## 6. If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades? 

```{r}
deck <- cards(jokers = FALSE) %>%
  mutate(black_card = str_detect(suit, "Club|Spade"))

n_black_cards <- deck %>%
  filter(black_card == TRUE) %>%
  summarise(n = n()) %>%
  pull(n)

deck %>%
  filter(suit == "Spade") %>%
  summarise(p = n() / n_black_cards)
```

