---
title: "Confidence Intervals - Homework"
output: html_notebook
---

# MVP
Now we’ll go back to CI creation in the normal fashion. We’ll take the ames data from the CIs lab earlier today and regard it now as a sample, we won’t be drawing any smaller samples from within it. This is the usual situation in an analysis: you use all the data available to you!

```{r}
library(tidyverse)
library(janitor)
library(infer)
```



## Task 1.
Load the data again, clean_names(), and re-familiarise yourself with it
```{r}
ames <- read_csv("data/ames.csv") %>%
  clean_names()

head(ames)
names(ames)
```


## Task 2.
Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?
```{r}
ames %>%
  ggplot() +
  aes(x = lot_area) +
  geom_histogram(aes(y = ..density..), col = "white", bins = 30) +
  stat_function(
    fun = dnorm,
    args = list(
      mean = mean(ames$lot_area),
      sd = sd(ames$lot_area)
    ),
    col = "red"
  )

ames %>%
  ggplot() +
  aes(x = lot_area) +
  geom_boxplot()
```
The distribution here almost looks normal to begin with, the majority of lot_area values are less than 25000, but it is skewed right by a relatively small amount of values much greater than this. It may look more normal without the outliers


## Task 3.
Compute and visualise a bootstrap sampling distribution for the mean(lot_area) of the sold houses.
```{r}
# generate bootstrap with 5000 reps, calculate mean(lot_area) for each rep
bootstrap_lot_area <- ames %>%
  specify(response = lot_area) %>%
  generate(reps = 5000, "bootstrap") %>%
  calculate(stat = "mean")

# plot sampling distribution
bootstrap_lot_area %>%
  ggplot() +
  aes(x = stat) +
  geom_histogram(col = "white")
```

This distribution looks much better.


## Task 4.
Use your bootstrap distribution to calculate a 95% CI for mean(lot_area), and visualise it on the distribution
```{r}
# use get_ci function to find 95% CI
lot_area_ci_95 <- bootstrap_lot_area %>%
  get_ci(level = 0.95, type = "percentile")

lot_area_ci_95

# use visualise() and shade_ci() to visualise CI on plot
bootstrap_lot_area %>%
  visualise(bins = 30) +
  shade_ci(endpoints = lot_area_ci_95)
```


## Task 5.
You would like to know the mean(lot_area) of the sold houses with higher confidence. Calculate the 99% CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the 95% CI? Does that make sense?
```{r}
# use get_ci function to find 99% CI
lot_area_ci_99 <- bootstrap_lot_area %>%
  get_ci(level = 0.99, type = "percentile")

lot_area_ci_99

# use visualise() and shade_ci() to visualise CI on plot
bootstrap_lot_area %>%
  visualise(bins = 30) +
  shade_ci(endpoints = lot_area_ci_99)
```
The 99% CI is slightly wider than the 95% CI. This makes sense as we are increasing the level of confidence in the interval, therefore including more mean values from our sampling distribution.


## Task 6.
Calculate the point estimate of the mean(lot_area)
```{r}
point_estimate <- bootstrap_lot_area %>%
  summarise(mean_lot_area = mean(stat))

point_estimate
```

# Extension

## Task 1.
Calculate a point estimate and 95% CI for the proportion of houses in the data built before 1920. Does the number of reps you use matter? [Investigate reps from 200 up to 50000, memory of your laptop permitting]. 

```{r}
# First create TRUE/FALSE variable for houses based on year_built < 1920
ames_before_1920 <- ames %>%
  mutate(built_before_1920 = if_else(year_built < 1920, TRUE, FALSE))

ames_before_1920 %>%
  group_by(built_before_1920) %>%
  summarise(count = n())
```


```{r}
# Bootstrap for 200 reps
bootstrap_year_built_200 <- ames_before_1920 %>%
  specify(response = built_before_1920, success = "TRUE") %>%
  generate(reps = 200, type = "bootstrap") %>%
  calculate(stat = "prop") 

# calculate 95% CI
ci_200_reps <- bootstrap_year_built_200 %>%
  get_ci(level = 0.95, type = "percentile")
  
# create summary
summary_bootstrap_200 <- bootstrap_year_built_200 %>%
  summarise(point_est = mean(stat)) %>%
  bind_cols(ci_200_reps)

summary_bootstrap_200
```

```{r}
# Bootstrap for 1000 reps
bootstrap_year_built_1000 <- ames_before_1920 %>%
  specify(response = built_before_1920, success = "TRUE") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") 

# calculate 95% CI
ci_1000_reps <- bootstrap_year_built_1000 %>%
  get_ci(level = 0.95, type = "percentile")
  
# create summary
summary_bootstrap_1000 <- bootstrap_year_built_1000 %>%
  summarise(point_est = mean(stat)) %>%
  bind_cols(ci_1000_reps)

summary_bootstrap_1000
```

```{r}
# Bootstrap for 10000 reps
bootstrap_year_built_10000 <- ames_before_1920 %>%
  specify(response = built_before_1920, success = "TRUE") %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "prop") 

# calculate 95% CI
ci_10000_reps <- bootstrap_year_built_10000 %>%
  get_ci(level = 0.95, type = "percentile")
  
# create summary
summary_bootstrap_10000 <- bootstrap_year_built_10000 %>%
  summarise(point_est = mean(stat)) %>%
  bind_cols(ci_10000_reps)

summary_bootstrap_10000
```

```{r}
# Bootstrap for 50000 reps
bootstrap_year_built_50000 <- ames_before_1920 %>%
  specify(response = built_before_1920, success = "TRUE") %>%
  generate(reps = 50000, type = "bootstrap") %>%
  calculate(stat = "prop") 

# calculate 95% CI
ci_50000_reps <- bootstrap_year_built_50000 %>%
  get_ci(level = 0.95, type = "percentile")
  
# create summary
summary_bootstrap_50000 <- bootstrap_year_built_50000 %>%
  summarise(point_est = mean(stat)) %>%
  bind_cols(ci_50000_reps)

summary_bootstrap_50000
```

            point_est   lower_ci    upper_ci
200 reps:   0.05990444	0.0518686   0.06826792	
1000 reps:  0.05976758	0.05153584	0.06791809
10000 reps: 0.06002765	0.05153584	0.06894198	
50000 reps: 0.06005062	0.05153584	0.06894198	

The amount of reps does not seem to make a signicant difference on the point estimate and CI on a the proportion.